---
title: 'Advanced R: Chapter 17'
output: html_document
date: "2023-06-27"
editor_options: 
  chunk_output_type: console
---

# 17.2

```{r}
library(rlang)
library(lobstr)
```

`expr()` captures whatever you write

```{r}
expr(mean(x, na.rm = TRUE))
expr(10 + 100 + 1000)
```

But it doesn't work if you use it in a function

```{r}
capture_it <- function(x) {
  expr(x)
}
capture_it(a + b + c)
```

captures "x", not "a + b + c", what the user supplied

for that, we use `enexpr()`

```{r}
capture_it <- function(x) {
  enexpr(x)
}
capture_it(a + b + c)
```

the `en` means "enrich". This is the "lazy-evaluated" form of expr. "Lazy" 
because it "waits" until it needs to be evaluated, rather than immediately after
being defined (previous example).

New terminology: `enexpr` "quotes" it's argument

Expressions are like lists

```{r}
f <- expr(f(x = 1, y = 2))
f

# Add a new argument
f$z <- 3
f

# Or remove an argument:
f[[2]] <- NULL
f
```

first element of an expression (or "call") is the function itself. 

# 17.3 Call = tree

Code is almost always represented as an abstract syntax tree (AST)

`lobstr::ast()` displays the AST of an expression

```{r}
lobstr::ast(f(a, "b"))
```

`a` is a symbol
`"b"` is a constant (string)

```{r}
lobstr::ast(f1(f2(a, b), f3(1, f4(2))))
```

# 17.4 code generates code

e.g. `rlang::call2` generates a function call from it's arguments:

```{r}
call2("f", 1, 2, 3)
call2("+", 1, call2("*", 2, 3))
```

alternatively, bang bang (aka unquote) operator can construct function calls too:

```{r}
xx <- expr(x + x)
yy <- expr(y + y)

expr(!!xx / !!yy)
```

This allows defining functions that can construct functions based on user input

```{r}
cv <- function(var) {
  var <- enexpr(var) # quote user expression
  expr(sd(!!var) / mean(!!var)) # insert into expression
}

cv(x)
cv(x + y)
```

# 17.5 eval runs code

`base::eval()` takes an expression and an environment and evaluates

```{r}
eval(expr(x + y), env(x = 1, y = 10))
eval(expr(x + y), env(x = 2, y = 100))
```

if env omitted, takes global env

```{r}
x <- 10
y <- 100
eval(expr(x + y))
```

# 17.6 customize expressions with functions

environmental varibales can be functions

here, operators are overwritten

```{r}
string_math <- function(x) {
  e <- env(
    caller_env(),
    `+` = function(x, y) paste0(x, y),
    `*` = function(x, y) strrep(x, y)
  )

  eval(enexpr(x), e)
}

name <- "Hadley"
string_math("Hello " + name)
string_math(("x" * 2 + "-y") * 3)
```


# 17.7 customize expressions with data

eval_tidy = base::eval

except eval_tidy uses dataframe as datamask (environment)
- where variables are columns


```{r}
df <- data.frame(x = 1:5, y = sample(5))
eval_tidy(expr(x + y), df)
```

with `enexpr` to use in a function:

```{r}
with2 <- function(df, expr) {
  eval_tidy(enexpr(expr), df)
}

with2(df, x + y)
```

# 17.8 quosures

address this problem

```{r}
with2 <- function(df, expr) {
  a <- 1000
  eval_tidy(enexpr(expr), df)
}
df <- data.frame(x = 1:3)
a <- 10
with2(df, x + a)
```

We want `with2`to evaluate using a defined in global (a=1000) not within with2 
definition (a=10).

for this we can use a quosure, which bundles an expression with an environment

```{r}
with2 <- function(df, expr) {
  a <- 1000
  eval_tidy(enquo(expr), df)
}

with2(df, x + a)
```

Where the environment is global and the expression is global

So for a datamask, always use enquo instead of enexpr
