---
title: "adv-r_ch18"
output: md_document
date: "2023-06-27"
editor_options: 
  chunk_output_type: console
---

```{r}
library(rlang)
library(lobstr)
```

# 18.2.4 exercises

AST to code: 

```{r}
#> █─f 
#> └─█─g 
#>   └─█─h
#> █─`+` 
#> ├─█─`+` 
#> │ ├─1 
#> │ └─2 
#> └─3
#> █─`*` 
#> ├─█─`(` 
#> │ └─█─`+` 
#> │   ├─x 
#> │   └─y 
#> └─z

ast(
  f(g(h()))
)
ast(
  (1+2)+3
)
ast(
  (x + y ) * z
)
```

# 18.3.2 symbols

create symbol

```{r}
expr(x) # raw
rlang::sym('x') # chr to symbol
```

symbol to string

```{r}
as.character(expr(x))
as_string(expr(x))
```

symbols cannot be vectorized
but multiple can be stored in a list (with `rlang::syms()`)

# 18.4 parsing and grammar

operator precedence for example: 1 + 2 * 3 = (1 + 2) * 3, or 1 + (2 * 3)

! has very low operator precedence, allowing expressions such as

```{r}
lobstr::ast(!x %in% y)
```

associativity is when repeated usage of infix operator results in ambiguity

is  1 + 2 + 3 = (1 + 2) + 3 or 1 + (2 + 3), they are equivalent in this case

but for example, ggplot2 + layers are non-associative 

e.g. geom_smooth + geom_point is not the same plot as geom_point + geom_smooth

most operators are left-associative, operations are evaluaetd from left to right

# 18.4.3 parsing and deparsing

`rlang::parse_expr` converts a string into an expression 

```{r}
x1 <- "y <- x + 10"
x1
is.call(x1)

x2 <- rlang::parse_expr(x1)
x2
is.call(x2)
```

deparsing: expr -> string

```{r}
z <- expr(y <- x + 10)
expr_text(z)
```

# 18.5 AST with recursive functions

```{r}
expr_type <- function(x) {
  if (rlang::is_syntactic_literal(x)) {
    "constant"
  } else if (is.symbol(x)) {
    "symbol"
  } else if (is.call(x)) {
    "call"
  } else if (is.pairlist(x)) {
    "pairlist"
  } else {
    typeof(x)
  }
}

expr_type(expr("a"))
expr_type(expr(x))
expr_type(expr(f(1, 2)))
```

```{r}
switch_expr <- function(x, ...) {
  switch(expr_type(x),
    ...,
    stop("Don't know how to handle type ", typeof(x), call. = FALSE)
  )
}
```

```{r}
recurse_call <- function(x) {
  switch_expr(x,
    # Base cases
    symbol = ,
    constant = ,

    # Recursive cases
    call = ,
    pairlist =
  )
}
```


